---
title: "Processing Code"
---

To start, I'll load packages. As I need packages along the way, I'll return here and add them.
```{r}
library(tidyverse)
library(gt)
library(webshot2)
library(here)
```

Now, loading data using the here package.
```{r}
here() #set working directory
data_original <- readRDS(here("data","raw-data", "NHIS_COVID.rds"))
data1 <- labelled::remove_val_labels(data_original) #removes labels from haven package which cause problems
```

Selecting variables we need will make the dataset smaller and quicker to work with. Note it is renamed at each step for reproducibility. 
```{r}
names(data_original) #view names of variables
data1 <- data_original %>%
  select(STRATA, PSU, PERNUM, SAMPWEIGHT, AGE, PA18AERSTR, HRSLEEP, DEPFREQ, PHQSLEEP, CMPSUPPORT, CVDTESTRSLT)
```


In the following code chunk, exploration and definition of each of these variables will be completed. These definitions and unique values help me to know how to define my binary vairables in the next step. Several of these vairable have codes for unknown, don't know, etc. I will need to convert these to NA.
```{r}
unique(data1$AGE)
unique(data1$PA18AERSTR)
unique(data1$HRSLEEP)
unique(data1$DEPFREQ)
unique(data1$PHQSLEEP)
unique(data1$CMPSUPPORT)
unique(data1$CVDTESTRSLT)
```


We will create binary vairables in a new dataset. For each variable, those who answered NA, 0, 996, 997, 998, 999 will be converted to NA for the remainder of that variable for the continuation of the analysis.
```{r}
data2 <- data1 %>%
  mutate(
    PA18AERSTR = ifelse(PA18AERSTR %in% c(0, 7, 8, 9), NA, PA18AERSTR), #put NA, 0, 6, 7, 8, 9 to NA
    EXERCISE = if_else(PA18AERSTR %in% c(2,3,4), 1, 0, missing = NA_real_) #create EXERCISE variable, 1 if guidelines by Am Heart Assoc met for aerobics, strength, or both, 0 if not met
    ) %>%
    mutate(
    AGE = ifelse(AGE %in% c(997, 999), NA, AGE),  #invalid ages to NA
    SLEEP = case_when( #Based on age, amount of sleep required from https://www.nhlbi.nih.gov/health/sleep/how-much-sleep#:~:text=Experts%20recommend%20that%20adults%20sleep,or%20more%20hours%20a%20night.
      AGE < 1  ~ if_else(HRSLEEP >= 12 & HRSLEEP <= 16, 1, 0, missing = NA_real_),
      AGE >= 1 & AGE < 3       ~ if_else(HRSLEEP >= 11 & HRSLEEP <= 14, 1, 0, missing = NA_real_),
      AGE >= 3 & AGE < 6       ~ if_else(HRSLEEP >= 10 & HRSLEEP <= 13, 1, 0, missing = NA_real_),
      AGE >= 6 & AGE < 13      ~ if_else(HRSLEEP >= 9 & HRSLEEP <= 12, 1, 0, missing = NA_real_),
      AGE >= 13 & AGE <= 18    ~ if_else(HRSLEEP >= 8 & HRSLEEP <= 10, 1, 0, missing = NA_real_),
      TRUE                     ~ NA_real_  #default NA for unspecified ages
    )) %>%
  mutate(
    DEPRESSED = case_when(
      DEPFREQ %in% c(1, 2) ~ 1, #daily and weekly is considered depressed
      DEPFREQ %in% c(3, 4, 5)    ~ 0, #monthly and a few times a year and never are considered not depressed
      DEPFREQ %in% c(0, 7, 8, 9) ~ NA_real_,#NIU & Unknowns are NA
      TRUE ~ NA_real_ #default NA for unexpected values
    )) %>%
  mutate(
    TROUBLE_SLEEPING = case_when(
      PHQSLEEP %in% c(1, 2, 3) ~ 1, #several days, more than half days, nearly every day are trouble sleeping
      PHQSLEEP == 0            ~ 0,  #not at all is no trouble sleeping
      PHQSLEEP %in% c(6, 7, 8, 9) ~ NA_real_,  #NIU & Unknowns are NA
      TRUE ~ NA_real_  #default NA for unexpected values
    )) %>%
  mutate(
    SOCIAL = case_when(
      CMPSUPPORT == 1 ~ 0,   #more social and emotional support
      CMPSUPPORT == 2 ~ 1,  #less social and emotional support
      CMPSUPPORT == 3 ~ 0,   #about the same
      CMPSUPPORT %in% c(0, 7, 8, 9) ~ NA_real_,  # NIU & Unknowns → NA
      TRUE ~ NA_real_  #default NA for unexpected values
    )) %>%
  mutate(
    COVID = case_when(
      CVDTESTRSLT == 2 ~ 1,  #yes, had COVID
      CVDTESTRSLT == 1 ~ 0,  #no, did not have COVID
      CVDTESTRSLT %in% c(0, 3, 7, 8, 9) ~ NA_real_,  #NIU, did not receive results, unknowns, and NA are NA
      TRUE ~ NA_real_  #default NA for unexpected values
    ))
  
```

For our final step, we will save this data to the processed_data folder.
```{r}
#save data2 as data3 in the processed-data folder
here()
write_rds(data2, here("data","processed-data", "data3.rds"))
```

